---
title: "SalpPOOP Project Sample Pre-Processing"
author: "Hannah Budroe"
date: "`r Sys.Date()`"
output: html_document
---

This follows first the Pre-processing script, then the phyloseq prep scripts, and any other additional processing found in analysis scripts

#Libraries
library(googlesheets4)
#library(plyr)
library(tidyr)
library(stringr)
library(dplyr)
library(ggplot2)
library(stats)
#install.packages("factoextra")
library(factoextra)
library(tibble)
library(phyloseq)
library(vegan)
library(data.table)

###Process environmental metadata ----

Collate biomass data + collection data
```{r}
#set directory for google sheet
googdata="https://docs.google.com/spreadsheets/d/1tKhRrnsyRMNwpQpYlr_9HUDoBUYpVK4RIApvrt_piFU/edit#gid=0"

#read biomass + filters into R from google sheets
ZoopBiomass= read_sheet(googdata)
FilterWeights=read_sheet(googdata,sheet=2)

#merge biomass + filters -> join by filter + fill in blanks
Biomass=plyr::join(ZoopBiomass,FilterWeights, type="left") %>% tidyr::fill(Station,Tow, Net)

#Subtract filter weights + group together samples on multiple filters
Biomass$FinalDW=Biomass$TotalDW-Biomass$Weight
dplyr::group_by(Biomass,Size,MinSize,Net,Tow,Station) %>%
  dplyr::summarise(FinalizedWeight=sum(FinalDW)) %>%
  dplyr::arrange(Station, Tow, Net, Size,MinSize,FinalizedWeight)-> Biomass

#read in tow specs, depths, lat/long/splits + adjust for desired data
DepthData=read_sheet(googdata, sheet=3) %>% tidyr::fill(Tow,Site,Station)
TowSpecs=read_sheet(googdata,sheet=4)
StationLog=read_sheet(googdata,sheet=5)
StationLog=StationLog[,c(1,26:27)]
Splits=read_sheet(googdata,sheet=6)
Splits$Size=str_sub(Splits$Size, end=-3) #truncate mm off size fract
Splits=fill(Splits,Volume_Filtered) #fill in vol filtered 

#merge everything together
Biomass <- plyr::join(Biomass,DepthData, type="left") %>% plyr::join(TowSpecs,type="left") %>%
  plyr::join(StationLog, type="left") %>% plyr::join(Splits,type="left")

#Normalize the biomass data
Biomass$Splits=as.numeric(Biomass$Splits)
Biomass$NormalWeight=(Biomass$FinalizedWeight/Biomass$Volume_Filtered)/Biomass$Splits
```
Note to self, the $NormalWeight biomass data is not the final value

Add in MOCNESS CTD data and merge with biomass data
```{r}
###add in MOCNESS CTD data
MOCNESS=read_sheet(googdata,sheet=7) %>% fill(Station)
#only include averaged columns
indx <- grepl('avg', colnames(MOCNESS))
MOCNESS=MOCNESS[indx] %>% cbind(MOCNESS[,c(1:2)])

#change mocness data col names
colnames(MOCNESS)=c("Pressure","Temperature","PotentialTemperature","Salinity",
                    "Density","Transmissivity","Fluorescence","Oxygen","Angle",
                    "HorizontalVelocity","VerticalVelocity","Station","Net")
#fix the net naming discrepancy between ctd and biomass -- corrected 0:8 -> 1-9
MOCNESS1<-MOCNESS
MOCNESS$Net=MOCNESS$Net+1

#duplicate station 282 ->286 --- stat286 data was lost, so use 282 similar data
dup282=MOCNESS[MOCNESS$Station == 282, ]
dup282$Station=286
MOCNESS=rbind(MOCNESS,dup282)

#join with biomass
Biomass=plyr::join(Biomass,MOCNESS, type="left")

#fix biomass units by converting g-> mg
Biomass$BiomassmgCm3 <- Biomass$NormalWeight*1000

#add in a few extra vars for classification
#add in epi v meso
Biomass$DepthZone <- ifelse(Biomass$MaxDepth <=200,"Epipelagic","Mesopelagic")
#a column just with cycle (no day)
Biomass$Cycle=substr(Biomass$`Cycle/Day`,start=1,stop=2)
```
For now, $BiomassmgCm3 is the final biomass value since it has correct units

(move to phyloseq prep script)

Clean up the sample names in biomass data, clean up cols, add ID vars
```{r}
###clean up the sample data
#in sampdata, make sure that the condensed biomass samples are reflected
sampdata<-Biomass
sampdata$test <-ifelse(sampdata$MinSize>=1,"Group","NoGroup") #select out groupings for condensing
sampdata$test2 <- ifelse(sampdata$MinSize<1,sampdata$MinSize,"x")
sampdata$test3=paste0(sampdata$test,sampdata$test2,sampdata$Net,sampdata$Tow) #create column to group by
sampdata %>% dplyr::group_by(test3) %>% summarise(condensedBiomass=sum(BiomassmgCm3)) ->testdf
#the summarised data is shorter length than the original data so need to duplicate somehow
left_join(sampdata,testdf,by="test3")->sampdata #it worked! now clean up

#take out pressure,volfiltered, the angles, finalizedweight, date, velocities, split/splits, test columns to condense biomass
sampdata=sampdata[,-c(11:13,19:23,31:33,37:39)] 
#create better sample ID
sampID=paste0(Biomass$`Cycle/Day`, "_T",Biomass$Tow, "_N",Biomass$Net,"_",Biomass$MinSize)
sampdata$sampID <-sampID
#index meta by sample IDs
sampdata=column_to_rownames(sampdata,var="sampID")

#add water mass cycle descriptor 
sampdata$ID <- ifelse(sampdata$WaterType=="Subantarctic",paste0("SA",substr(sampdata$`Cycle`,start=2,stop=2)),paste0("ST",substr(sampdata$`Cycle`,start=2,stop=2)))
#8/19/23 realized made mistake and fix nomenclature
sampdata$ID[sampdata$ID == 'SA5'] <- 'SA3'
sampdata$ID[sampdata$ID == 'ST3'] <- 'ST1'
sampdata$ID[sampdata$ID == 'ST4'] <- 'ST2'
```

