---
title: "SalpPOOP Project Analyses 9_20_24"
author: "Hannah Budroe"
date: "2024-09-20"
output: html_document
---

#This script is intended for analyses of mesozooplankton diversity and community structure ~salp bloom ~size fractions ~depth ~water masses
#Script created after (new and final) Github repository was created [I will be better at version control from now on]
#many of the initial version analyses will be imported from Rmd scripts I previously created (SalpPOOP Paper Figures V2, etc.) & updated

#Libraries ----
library(tidyverse)
library(vegan)
library(phyloseq)

#Functions ----

###First need to output required files to new computer
Looking for mzg (and pr2) phylo objects <-
(see salpPOOPthesis_finalplots_scriptR/phyloseq prep for final processing 5/18/23)
want salpPOOP_mzg_meta/otu/tax
```{r}
#here I'm getting stats for the initally created, unfiltered mzgps object
mzgps #12771 taxa, 196 samples, 9 tax ranks
sum(sample_sums(mzgps)) #10006100
sp_accum_mzgps <- specaccum(t(otu_table(mzgps)), method = "random")
#plot(sp_accum_mzgps, col = "cadetblue1", ylab = "ASVs (10^3)",
     xlab = "#Samples", ci.type = "poly", lwd=2, ylim = c(1, length(t(otu_table(mzgps)))+1000), ci.lty = 0, ci.col = alpha("cadetblue1", alpha = 0.2), cex.axis=0.6) #doesn't plot - figure out later?
```


###Section 1 () ----

###Figure 2 (Compositional Depth Plots) ----
#Figure 2a: Total community read/biomass abundance ~depth
```{r}
#add in columns (interval width + depth midpoints) to physeq object
sample_data(mzg)$BinWidth = sample_data(mzg)$MaxDepth - sample_data(mzg)$MinDepth
sample_data(mzg)$Midpoint = (sample_data(mzg)$BinWidth/2) + sample_data(mzg)$MinDepth

#Combine size fraction biomass, prune for ease of plotting right now
mzg %>% sample_data() %>% dplyr::group_by(Net,Tow,Station) %>% dplyr::summarise(Biomass=sum(condensedBiomass)) ->mdf1 #df of added biomass
mzg %>% filter_taxa(function(x) sum(x) > 30 , TRUE) %>% psmelt() -> mdf
#mzg %>% psmelt() -> mdf
#prune taxa with less >30reads

mdf1 %>% right_join(mdf) ->mdf
disp1 <- max(mdf$Abundance)*3.3
mdf$Abundance[mdf$DielPattern == "Night"] <- mdf$Abundance[mdf$DielPattern == "Night"]*-1 
mdf$Biomass[mdf$DielPattern == "Night"] <- mdf$Biomass[mdf$DielPattern == "Night"]*-1

(p <- ggplot(mdf, aes(width = BinWidth*.9)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin =  -Inf, ymax = 0, fill = "grey60", alpha = 0.5) +
  geom_bar(aes(x = Midpoint, y = Abundance, fill = Order), stat = "identity") +
  geom_point(aes(x = Midpoint, y = Biomass*3000,group=DielPattern)) +
  geom_line(aes(x = Midpoint, y = Biomass*3000, group = DielPattern)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 4), limits = c(-disp1, disp1), labels=abs, sec.axis = sec_axis(~./3000, name = bquote('MOCNESS Biomass '(mgC/m^3)), scales::pretty_breaks(n = 4), labels = abs)) + #define the name of your second axis
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90)) +
  #scale_fill_brewer(type = "qual", palette = "Set1") + #fill for each species is different, using a rcolorbrewer palette called "Set1"
  scale_color_manual(values = "black", name = paste('Zooscan biomass', sep = '')) +# name = bquote(atop('Zooscan Biomass ', (mgC/m^2)))) +  #color for the zooscan biomass is black
  coord_flip() + 
  scale_x_reverse() + 
  labs(x = "Depth (m)", y = "Relative abundance of 18S sequences") + 
  geom_hline(yintercept = 0, linewidth = 0.2)+
  facet_grid(~ID, scales="free_x"))
```
#Figure 2b: Order == Calanoid read abundance ~depth 
```{r}
#add col for vertical bin sizes (depth) + midpoint (abundance line) - doesn't change with tax
sample_data(mzg)$BinWidth = sample_data(mzg)$MaxDepth - sample_data(mzg)$MinDepth
sample_data(mzg)$Midpoint = (sample_data(mzg)$BinWidth/2) + sample_data(mzg)$MinDepth

mzg %>% sample_data() %>% dplyr::group_by(Net,Tow,Station) %>% dplyr::summarise(Biomass=sum(condensedBiomass)) ->mdf1 #df of added biomass
mzg %>% subset_taxa(Order=="Calanoida") %>% psmelt() -> mdf_calanoida
mdf1 %>% right_join(mdf_calanoida) ->mdf

#skip the biomass summarization, mdf joining with mdf1 ?
disp1 <- max(mdf$Abundance)*3.3 
mdf$Abundance[mdf$DielPattern == "Night"] <- mdf$Abundance[mdf$DielPattern == "Night"]*-1 
mdf$Biomass[mdf$DielPattern == "Night"] <- mdf$Biomass[mdf$DielPattern == "Night"]*-1

#Plot
(p <- ggplot(mdf, aes(width = BinWidth*.9)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin =  -Inf, ymax = 0, fill = "grey60", alpha = 0.5) +
  geom_bar(aes(x = Midpoint, y = Abundance, fill = Family), stat = "identity") +
  geom_point(aes(x = Midpoint, y = Biomass*3000,group=DielPattern)) +
  geom_line(aes(x = Midpoint, y = Biomass*3000, group = DielPattern)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 4), limits = c(-disp1, disp1), labels=abs, sec.axis = sec_axis(~./3000, name = bquote('Total MOCNESS Biomass '(mgC/m^3)), scales::pretty_breaks(n = 4), labels = abs)) + #define the name of your second axis
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90)) +
  #scale_fill_brewer(type = "qual", palette = "Set1") + #fill for each species is different, using a rcolorbrewer palette called "Set1"
  scale_color_manual(values = "black", name = paste('Zooscan biomass', sep = '')) +# name = bquote(atop('Zooscan Biomass ', (mgC/m^2)))) +  #color for the zooscan biomass is black
  coord_flip() + 
  scale_x_reverse() + 
  labs(x = "Depth (m)", y = "Relative abundance of 18S Calanoida sequences") + 
  geom_hline(yintercept = 0, linewidth = 0.2)+
  facet_grid(~ID, scales="free_x"))
```
#Figure 2c: Class == Malacostraca read abundance ~depth
```{r}
#add col for vertical bin sizes (depth) + midpoint (abundance line) - doesn't change with tax
sample_data(mzg)$BinWidth = sample_data(mzg)$MaxDepth - sample_data(mzg)$MinDepth
sample_data(mzg)$Midpoint = (sample_data(mzg)$BinWidth/2) + sample_data(mzg)$MinDepth

mzg %>% sample_data() %>% dplyr::group_by(Net,Tow,Station) %>% dplyr::summarise(Biomass=sum(condensedBiomass)) ->mdf1 #df of added biomass
mzg %>% subset_taxa(Class=="Malacostraca") %>% psmelt() -> mdf_malacostraca
mdf1 %>% right_join(mdf_malacostraca) ->mdf

#skip the biomass summarization, mdf joining with mdf1 ?
disp1 <- max(mdf$Abundance)*3.3 
mdf$Abundance[mdf$DielPattern == "Night"] <- mdf$Abundance[mdf$DielPattern == "Night"]*-1 
mdf$Biomass[mdf$DielPattern == "Night"] <- mdf$Biomass[mdf$DielPattern == "Night"]*-1

#Plot
(p <- ggplot(mdf, aes(width = BinWidth*.9)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin =  -Inf, ymax = 0, fill = "grey60", alpha = 0.5) +
  geom_bar(aes(x = Midpoint, y = Abundance, fill = Order), stat = "identity") +
  geom_point(aes(x = Midpoint, y = Biomass*3000,group=DielPattern)) +
  geom_line(aes(x = Midpoint, y = Biomass*3000, group = DielPattern)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 4), limits = c(-disp1, disp1), labels=abs, sec.axis = sec_axis(~./3000, name = bquote('Total MOCNESS Biomass '(mgC/m^3)), scales::pretty_breaks(n = 4), labels = abs)) + #define the name of your second axis
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90)) +
  #scale_fill_brewer(type = "qual", palette = "Set1") + #fill for each species is different, using a rcolorbrewer palette called "Set1"
  scale_color_manual(values = "black", name = paste('Zooscan biomass', sep = '')) +# name = bquote(atop('Zooscan Biomass ', (mgC/m^2)))) +  #color for the zooscan biomass is black
  coord_flip() + 
  scale_x_reverse() + 
  labs(x = "Depth (m)", y = "Relative abundance of 18S Malacostracan sequences") + 
  geom_hline(yintercept = 0, linewidth = 0.2)+
  facet_grid(~ID, scales="free_x"))
```
#Figure 2d? All GZ read abundance ~depth
```{r}
#add col for vertical bin sizes (depth) + midpoint (abundance line) - doesn't change with tax
sample_data(mzg)$BinWidth = sample_data(mzg)$MaxDepth - sample_data(mzg)$MinDepth
sample_data(mzg)$Midpoint = (sample_data(mzg)$BinWidth/2) + sample_data(mzg)$MinDepth

mzg %>% sample_data() %>% dplyr::group_by(Net,Tow,Station) %>% dplyr::summarise(Biomass=sum(condensedBiomass)) ->mdf1 #df of added biomass
mzg %>% subset_taxa(Class == "Hydrozoa" | Class == "Appendicularia" | Class == "Scyphozoa" | Class == "Thaliacea" | Phylum == "Ctenophora") %>% psmelt() -> mdf_gz
mdf1 %>% right_join(mdf_gz) ->mdf

#skip the biomass summarization, mdf joining with mdf1 ?
disp1 <- max(mdf$Abundance)*3.3 
mdf$Abundance[mdf$DielPattern == "Night"] <- mdf$Abundance[mdf$DielPattern == "Night"]*-1 
mdf$Biomass[mdf$DielPattern == "Night"] <- mdf$Biomass[mdf$DielPattern == "Night"]*-1

#Plot
(p <- ggplot(mdf, aes(width = BinWidth*.9)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin =  -Inf, ymax = 0, fill = "grey60", alpha = 0.5) +
  geom_bar(aes(x = Midpoint, y = Abundance, fill = Order), stat = "identity") +
  geom_point(aes(x = Midpoint, y = Biomass*3000,group=DielPattern)) +
  geom_line(aes(x = Midpoint, y = Biomass*3000, group = DielPattern)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 4), limits = c(-disp1, disp1), labels=abs, sec.axis = sec_axis(~./3000, name = bquote('Total MOCNESS Biomass '(mgC/m^3)), scales::pretty_breaks(n = 4), labels = abs)) + #define the name of your second axis
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90)) +
  #scale_fill_brewer(type = "qual", palette = "Set1") + #fill for each species is different, using a rcolorbrewer palette called "Set1"
  scale_color_manual(values = "black", name = paste('Zooscan biomass', sep = '')) +# name = bquote(atop('Zooscan Biomass ', (mgC/m^2)))) +  #color for the zooscan biomass is black
  coord_flip() + 
  scale_x_reverse() + 
  labs(x = "Depth (m)", y = "Relative abundance of 18S GZ sequences") + 
  geom_hline(yintercept = 0, linewidth = 0.2)+
  facet_grid(~ID, scales="free_x"))
```


#Figure ? Alpha diversity 